---
title: "R Notebook"
output: html_notebook
---

ggplot2

#gg = grammar of graphics


to create a plot, call ggplot()


geoms = shape
aes = aesthetics

in the gpplot function, is where the + argument/function comes in (where you add layers)

fill inside the aes code, will set the fill colour, except if in inverted commas
code, then it will map it to the argugement, and if not in upside down commas,
then it will include the word hard coded in a legend

position = dodge will set an area next to a bar chart, to compare the areas

the dataset we're working with has 14 rows, but ggplot calculated this for us
automatically
to demonstate the quantity, do:
stat = count (as an argument)
 
perhaps better to do:
dplyr function
group_by(superpower, school_year)%>%
  summarise(n=n())
 
or
 dplyr function
 count(superpower, school_year)

# labelling

labs - is a function which can be used for labelling

labs(x = 'superpower',
      y = 'count',
      title = "preferred superpower by year",
      subtitle = "data from students",
      caption = "sourced from")

caption = goes in bottom right corner

they can also be put as seperate arguments in the function

```{r}
olympics_medals <- CodeClanData::olympics_overall_medals
```
```{r}
library(tidyverse)
```


```{r}
top_10 <- olympics_medals %>%
  filter(season == "Summer", medal == "Gold") %>%
  arrange(desc(count)) %>%
  top_n(10)
top_10
```

```{r}
top_10 %>%
  ggplot(aes(x = count, y = team)
         )
  
```

NOTE

To build the graphic:
- You cannot follow the documentation as aes argument MUST be INSIDE the aes
  argument, therefore you create the geoms BEFORE the aes, despite that being
  the FIRST layer
-Additionally, despite the argument being: geom_point
  This is CONVERTED when you set the type, e.g. column or bar
  
```{r}
ggplot(top_10) +
  geom_col(aes(x = count, y = team, fill = team)
         )
  
```

Layering of ggplot

Chickweight (is a dataset built into R)

```{r}
chick_weight <- janitor::clean_names(ChickWeight)
```

```{r}
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet)) 
```

adding another layer
```{r}
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet))  +
  geom_point(aes(x = time, y = weight))
```

adding geom_smooth, which adds an "average" line
```{r}
ggplot(chick_weight) +
  geom_line(aes(x = time, y = weight, group = chick, colour = diet))  +
  geom_point(aes(x = time, y = weight, colour = diet)) +
  geom_smooth(aes(x = time, y = weight, group = diet, colour = diet))
```

```{r}
students_superpower <- CodeClanData::students
```

```{r}
ggplot(students_superpower) +
  geom_point(aes(x = height_cm, y = reaction_time, colour = superpower))
```

```{r}
pets <- read_csv("pets.csv")
```
```{r}
ggplot(pets, aes(x = age, y = weight)) +
  geom_point(aes(colour = sex, shape = animal), size = 4) +
  geom_text(
    aes(label = name),
                nudge_x = 0.5,
                nudge_y = 0.1)

```

Saving reports
```{r}
last_plot()
#will repeat the last plot you made

ggsave()
# takes advantage of this function

# plots can also be assigned to variables



```

