---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
Scales and coordinates
Scales
Labels


by adding scale_x_continuous / scale_y, the plot is getting closer together
```{r}
ggplot(data = pets, mapping = aes (x = age, y = weight, colour = sex)) +
  geom_point() +
  scale_x_continuous(breaks = 1:13) +
  scale_y_continuous(breaks = c(5, 7, 9, 11))

```

```{r}
ggplot(data = pets, mapping = aes (x = age, y = weight, colour = sex)) +
  geom_point() +
  scale_x_continuous(breaks = 1:13, limits = c(1, 8)
                     )
```
The problem with limits is that if the dataset/line continues, and the limit is
then set at e.g. 10, the line will then chop off at 10, even though there is 
data beyond that (leaving a clear line, with data points on the x axis)
Coord_cartesian is a better way to zoom in on your dataset

```{r}
# d %>% 
#   ggplot(aes(x, y)) +
#   geom_path() +
#   coord_cartesian(xlim = c(1, 8))
```

TASK - eliminate pets who weigh more than 8kg from the plot
```{r}
ggplot(data = pets, mapping = aes(age, weight, colour = sex)) +
  geom_point() +
  scale_y_continuous(limits = c(1, 8))
```

```{r}
seq()
# creates a sequence of numbers
# e.g. # seq(12) will give 1 2 3 4 5 6 7 8 9 10 11 12
```

```{r}
ggplot(pets) +
  aes(age, weight, colour = sex) +
  geom_point() +
  scale_x_continuous(breaks = 1:13) # adjust breaks here
```

```{r}
ggplot(pets) +
  aes(age, weight, colour = sex) +
  geom_point() +
  scale_x_continuous(breaks = 1:13) + 
  scale_y_continuous(breaks = c(5, 7, 9, 11))
```

Flipping a plot

```{r}
ggplot(pets, mapping = aes(x = animal, fill= sex)) +
  geom_bar()
```
there are two ways to flip this plot
either swap which object is x and which is y

OR coord_flip() func

```{r}
ggplot(pets, mapping = aes(x = animal, fill= sex)) +
  geom_bar() +
  coord_flip()
```

Factors
```{r}
# factors are strings which have levels, but these are assigned levels, and these
# levels are assigned an alphabetical order
# factor levels are not intrinsically ordered (according to R)but they CAN
# be ordered, if assigned an ordered, if the argument for T is set
# you DON'T want factors to be ordered
# factors CAN be converted to numeric

```

Formatting of labels

```{r}
monthly_sales <- read_csv("monthly_sales.csv")
monthly_sales %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April"))) %>%
  ggplot(monthly_sales) +
  aes(x = branch, y = sales, fill = month) +
  geom_col(position = "dodge") +
  coord_flip()
```

```{r}
monthly_sales %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April")),
         sales_thousands = sales / 1000) %>%
  ggplot(aes(x = branch, y = sales_thousands, fill = month)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(y = "sales (1000 units)")
  
```

```{r}
monthly_sales %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April")),
         sales_thousands = sales / 1000) %>%
  ggplot(aes(x = branch, y = sales_thousands, fill = month)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::comma)+
  labs(y = "sales (1000 units)") +
  coord_flip()
```

log scales

```{r}
monthly_sales %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April")),
         sales_thousands = sales / 1000) %>%
  ggplot(aes(x = branch, y = sales_thousands, fill = month)) +
  geom_col(position = "dodge") +
  scale_y_log10() +
  coord_flip()
```

```{r}

```

