---
title: "R Notebook"
output: html_notebook
---

Lesson 2 - Facets

The plot with chickens is difficult because it shows too much information, but
by using an average of all of them together, you hide information, so you want to show as much
as possible.
```{r}
# this is where facet wrap comes in.

# it breaks up a graph into individual plots

```

facet_wrap()
```{r}
ggplot(chick_weight) + 
  aes(x = time, y = weight, group = chick, colour = diet) +
  geom_line() +
  facet_wrap(~ diet)
```

```{r}
# within facet wrap, you can specify nrow, ncol, scales
```



facet_grid
```{r}
ggplot(chick_weight) + 
  aes(x = time, y = weight, group = chick, colour = diet) +
  geom_line() +
  facet_wrap(~ diet, ncol = 4)

```


```{r}
unemployment <- read_csv("benefits.csv")
```

```{r}
Use geom_histogram to create a histogram showing the number of unemployed by age. Then facet the graph based on biological sex.
Hint: geom_histogram is almost the same as geom_bar, but is better suited to plotting continuous data because of the binning.
```

```{r}
ggplot(unemployment) + 
geom_histogram(aes(x = age, fill = sex), col = "white") + 
facet_wrap( ~ sex, ncol = 2)
```


```{r}
ggplot(unemployment) + 
  geom_histogram(aes(x = age, fill = sex), col = "white") + 
  facet_wrap(married ~ sex)
```

```{r}
ggplot(unemployment) + 
  geom_histogram(aes(x = age, fill = sex), col = "white") + 
  facet_grid(married ~ sex)
```


```{r}
# Task - 5 mins
#     Use the same dataset, but base the facet_grid() on marital status (married) and if the individual has children (dkids).
#     Extension: can you improve the way that the facets are labled.
```

```{r}
ggplot(unemployment) + 
  geom_histogram(aes(x = age, fill = dkids), col = "white") + 
  facet_grid(married ~ dkids)

ggplot(unemployment) + 
  geom_bar(aes(x = married, fill = dkids), col = "white") + 
  facet_grid(married ~ dkids, labeller = "label_value")

```

This misrepresents our data
```{r}
ggplot(unemployment) + 
  geom_bar(aes(x = married, fill = dkids), col = "white") + 
  facet_grid(married ~ dkids, scales = "free_y")
```

It is not terriblt useful on THIS dataset, because the names are short, but it
can be good on very long names of bar graphs as it "widens" the x axis
```{r}
ggplot(unemployment) + 
  geom_bar(aes(x = married, fill = dkids), col = "white") + 
  facet_grid(married ~ dkids, scales = "free_x")
```

