---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(ggplot2)
```


Colours and themes

Colours in R

3 types of colour scheme are generally used
sequential (logical order)
diverging(going in different order)
categorical

(all pretty self-explanatory)

R has its own inbuilt palette - palette()

ggplot has it own though

you can also use RGB encoding
```{r}
ggplot(pets, mapping = aes(weight, age)) +
  geom_point(colour = "#33ceff")
```

```{r}
ggplot(pets, aes(x = age, y = weight, colour = sleep)) +
  geom_point()+
  scale_colour_gradient(low = "grey5", high = "grey50")

ggplot(pets, aes(x = age, y = weight, colour = sleep)) +
  geom_point()+
  scale_colour_gradient(low = "red", high = "blue")

```

```{r}
ggplot(pets, aes(x = age, y = weight, colour = sleep)) +
  geom_point()+
  scale_colour_gradient(low = "light blue", high = "purple")
```

Can have a divergent scale
```{r}
ggplot(pets, aes(x = age, y = weight, colour = sleep)) +
  geom_point()+
  scale_colour_gradient2(midpoint = 10, low = "red", high = "purple")
```


```{r}
students_superpower %>%
  mutate(school_year = factor(school_year,
                              levels = c("Year 7",
                                         "Year 8",
                                         "Year 9",
                                         "Year 10",
                                         "Year 11",
                                         "Year 12"))) %>%
  ggplot(mapping = aes(x = school_year, fill = superpower)) +
  geom_bar() +
  scale_fill_manual(
    values = c("Fly" = "red",
                "Freeze time" = "blue",
                "Invisibility" = "white",
                "Telepathy" = "green")
                    )
```

improving colour choices with colour brewer

```{r}
ggplot(pets, aes(weight, age, colour = sleep)) +
  geom_point() +
  scale_colour_distiller(palette = "PiYG")
```

```{r}
ggplot(students_superpower, aes(x = school_year, fill = superpower)) +
  geom_bar() +
  scale_fill_brewer(palette = "PiYG")
```

Task - change the palette of students to a colour friendly one
```{r}
ggplot(students_superpower, aes(x = school_year, fill = superpower)) +
  geom_bar() +
  scale_fill_brewer(palette = "PuRd")
```

To set / develop / apply themes, you can use / install ggthemes
```{r}
# looks like a really old journal article!

ggplot(students_superpower, aes(x = school_year, fill = superpower)) +
  geom_bar() +
  theme_void() +
  facet_wrap(~superpower)
```

Themes available (just a short list)
```{r}
theme_minimal()
theme_bw()
theme_classic()
theme_light()
theme_void()
```

Theme elements
```{r}
# the grid
# the title
# labelling of axis

#by using theme() you can modify specific elements of the plot
# e.g.
# theme(axis.title = )
```

Task with custom theme
```{r}
scottish_exports <- read_csv("scottish_exports.csv")

ggplot(scottish_exports) +
  geom_line(aes(x = year, y = exports, colour = sector)) +
  facet_wrap(~sector, scales = 'free_y')
```

```{r}
my_theme <- 
```

