---
title: "R Notebook"
output: html_notebook
---

Week 11 Day 3 

Clustering Lab

Use k-means clustering to investigate potential relationships in the dataset students_big from the CodeClanData library.

#1 Part a
We’ve been asked to create clusters of students for the answers to the ‘importance’ questions (where students rate how important they find certain topics on a scale from 0 to 1000, the higher the number the more important they think the issue is).

If we are to use the mathematical methods (elbow etc.) what would your optimal value of k been? Do you think the data seems well suited to cluster?

```{r}
# reading in packages and data

library(tidyverse)
library(janitor)

students_data <- CodeClanData::students_big
```

Cleaning the data
```{r}
clean_student_data <- students_data %>%
  clean_names() %>%
  select(-country, -region, -gender, -ageyears, -handed, -height, -foot_length,
         -arm_span, -languages_spoken, -travel_to_school, -travel_time_to_school,
         -reaction_time, -score_in_memory_game, -favourite_physical_activity) %>%
  mutate_if(is.numeric, scale)
```

```{r}
clustered_student_data <- kmeans(clean_student_data,
                                 centers = 6,
                                 nstart = 25)

clustered_student_data
```

cleaning/accessing the data
```{r}
library(broom)

tidy(clustered_student_data,
     col.names = colnames(clean_student_data))
augment(clustered_student_data, clean_student_data)
```

building the nested tibble to draw the elbow graph / plot
```{r}
max_k <- 20

k_clusters_student_data <- tibble(k = 1:max_k) %>%
  mutate(kclust = map(k, ~ kmeans(clean_student_data, .x, nstart = 25)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, clean_student_data)
  )

k_clusters_student_data

```
unnesting it, to be able to plot it 
```{r}
clusterings_student_data <- k_clusters_student_data %>%
  unnest(glanced)
```

```{r}
ggplot(clusterings_student_data, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```

```{r}
library(factoextra)

fviz_nbclust(clean_student_data,
             kmeans,
             method = "silhouette",
             nstart = 25)
```


```{r}
fviz_nbclust(clean_student_data, 
             kmeans, 
             method = "gap_stat", 
             nstart = 25, 
             k.max = 10)
```

```{r}
fviz_nbclust(clean_student_data,
             kmeans,
             method = "wss",
             nstart = 25)
```

Probably 2.


## 2 Part b

We have been asked to cluster the variables which are asking about environmental questions (importance_reducing_pollution, importance_recycling_rubbish, importance_conserving_water, importance_saving_energy) into 2 clusters because the school wants to split the students into 2 groups to target with specific learning/lessons about environmental issues.


```{r}
clustered_student_data_1 <- kmeans(clean_student_data,
                                 centers = 1,
                                 nstart = 25)

clustered_student_data_1

clustered_student_data_2 <- kmeans(clean_student_data,
                                 centers = 1,
                                 nstart = 25)

clustered_student_data_2
```



Once you have set up the 2 clusters visualise a combination of the variables split by the clusters (e.g. plot importance_reducing_pollution vs importance_recycling_rubbish etc.) how might you describe these 2 groups in terms of characteristics on views on environmental issues back to the school?

```{r}
library(GGally)

clusterings_student_data %>% 
  unnest(cols = c(augmented)) %>%
  select(-iter, -betweenss, -tot.withinss, -totss, -k) %>%
  ggcorr()
```
```{r}
clusterings_student_data %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 2) %>%
  ggplot(aes(x = importance_recycling_rubbish, y = importance_conserving_water, colour = .cluster)) +
  geom_point(aes(color = .cluster))
```





```{r}
clusterings_student_data %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 2) %>%
  ggplot(aes(x = importance_recycling_rubbish, y = importance_conserving_water, colour = .cluster)) +
  geom_point(aes(color = .cluster)) +
  facet_wrap(~ .cluster)
```


