---
title: "R Notebook"
output: html_notebook
---
Week 11 Day 2

Decision Trees

Several types of nodes

Root nodes (at the top)
Leaf nodes (at the bottom, where there is an outcome)
Intermediate/decision nodes

There are many metrics which can be used to split data, the one generally used in R is GINI
You can also use R2, or mean r2, or even RMSE

```{r}
library(tidyverse)
library(rpart)
library(rpart.plot)

thrones <- read_csv("data/character_data_S01-S08.csv")
```

```{r}
head(thrones)
```

```{r}
clean_thrones<- thrones %>% 
  
# Only keep variables of interest
  select(c(sex, religion, occupation, social_status, allegiance_last, allegiance_switched, dth_flag, featured_episode_count, prominence)) %>% 
  
# Convert to factor level
    mutate(sex = factor(sex, levels = c(1, 2, 9), labels = c("Male", "Female", "Unknown")),
    religion = factor(religion, levels = c(0, 1, 3, 4, 5, 6, 7, 9), labels = c("Great Stallion", "Lord of Light", "Faith of the Seven", "Old Gods", "Drowned God", "Many Faced God", "Other", "Unknown")),
    occupation = factor(occupation, levels = c(1, 2, 9), labels = c("Silk-collar", "Leather-collar", "Unknown")),
    social_status = factor(social_status, levels = c(1, 2, 9), labels = c("Highborn", "Lowborn", "Unknown")),
    allegiance_last = factor(allegiance_last, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9), labels = c("Stark", "Targaryen", "Night's Watch", "Lannister", "Greyjoy", "Bolton", "Frey", "Other", "Unknown")),
    allegiance_switched = factor(allegiance_switched, levels = c(1, 2, 9), labels = c("No", "Yes", "Unknown")),
    dth_flag = factor(dth_flag, levels = c(0, 1), labels = c("Survived", "Died"))) %>%
  
# Remove NAs 
 na.omit()
```


```{r}
glimpse(clean_thrones)
```

FIrst step is to always create a test train split

## Test train split
```{r}
set.seed(19)
```

```{r}
n_data <- nrow(clean_thrones)
test_index <- sample(1:n_data, size = n_data*0.2)
thrones_test <- slice(clean_thrones, test_index)
thrones_train <- slice(clean_thrones, -test_index)
```

```{r}
thrones_test %>%
  janitor::tabyl(dth_flag)
```

```{r}
thrones_train %>%
  janitor::tabyl(dth_flag)
```

## Building the decision tree

```{r}
thrones_fit <- rpart(
  formula = dth_flag ~ .,
  data = thrones_train,
  method = 'class'
)

rpart.plot(thrones_fit,
           yesno = 2,
           fallen.leaves = TRUE,
           faclen = 6,
           digits = 4)

# digits is the number of digits displayed at the bottomg
# fallen.leaves is whether to display the completed nodes at the bottom of the graph

# in the first node, .58 is the chance of dying, 100% is how much data has been passed through that node
# all the probabilities are of dying
```

```{r}
higher_0.4812 <- thrones_train %>%
  filter(dth_flag == "Died") %>%
  nrow()

lower_0.4812 <- thrones_train %>%
  filter(prominence < 0.4812) %>%
  nrow()

92/(184+92)
```


```{r}
rpart.plot(thrones_fit,
           yesno = 2,
           fallen.leaves = TRUE,
           faclen = 6,
           digits = 4,
           type = 4,
           extra = 101)
```

```{r}
rules <- rpart.rules(thrones_fit, cover = TRUE)
```


Using the model to create a prediction

```{r}
library(modelr)

thrones_test_pred <- thrones_test %>%
  add_predictions(thrones_fit, type = "class")

thrones_test_pred %>%
  select(pred, dth_flag, prominence, religion, allegiance_last, featured_episode_count)
```

Building a confusion matrix
```{r}
library(yardstick)

conf_mat <- thrones_test_pred %>%
  conf_mat(truth = dth_flag, estimate = pred)

conf_mat
```

```{r}
accuracy <- thrones_test_pred %>%
  accuracy(truth = dth_flag, estimate = pred)
accuracy
```

Looking at the sensitivity and specificity
sens = TRUE positive rate
spec = TRUE negative rate

```{r}
thrones_test_pred %>%
  sensitivity(truth = dth_flag, estimate = pred)
```

```{r}
thrones_test_pred %>%
  specificity(truth = dth_flag, estimate = pred)
```

```{r}
library(caret)

confusionMatrix(thrones_test_pred$pred, thrones_test_pred$dth_flag)
```








