---
title: "Main Notebook"
output: html_notebook
---

## Document starts here
Louise Shambrook

Updated 11/11/2021

## 0.0 Index

0.1 - Libraries

0.2 - Data sources

0.3 - Class log

1.1 - Summary functions

1.2 - dplyr functions

1.3 - Other important functions

1.4 - str functions

1.5 - joining functions

1.5.1 - Pivot functions

1.6 - base R functions

1.6.1 - Useful mathematical operators

1.6.2 - Conditional operators

1.7 - Unit testing

2.1 - If statements

2.2 - For statements

2.3 - Which statements

2.4 - Functions

3.1 - Styling and best practice

3.2 - Reproducibility and file paths

3.3 - Assertive programming

4.1 - Tidy data

5.1 - Regular expressions (regex)

5.2 - Strings

6.1 - Matricies and lists

10.1 - Notes from weekend homework

## 0.1 Libraries

library(tidyverse)
library(readxl)
library(janitor)
library(styler)
library(fcuk)
library(errorist)
library(assertr)

## 0.2 Data Sources

library(CodeClanData)
data/dog_bites.csv
data/edinburgh_seedmix.xlsx
data/new_coders.csv
data/character-deaths.csv
mdata/movies.csv

To read in data

read_csv()

## 0.3 Class log

W1 D1 

W1 D2 - reading in excel files 

W1 D3 - 

W1 D4 - Recoding, Styling and best practice, errors

W1 D5 - 

W2 D1 - Joins and Tidy data

W2 D2 - Strings and Regular expressions

W2 D3 - Matrices and Lists, control flow, and functions

W2 D4 - Unit testing, assertive programming, reproducibility 

W2 D5 - 

## 1.1 Summary / data exploration

`nrow()` - returns the number of rows

`ncol()` - returns the number of columns

`dim()` - returns dimensions (rows and columns)

`names()` - returns the names of all variables / columns

`class()` - returns the type of an object

`head()` - returns the top 5 rows of a df (will accept a numerical arg (, 5))

`tail()` - returns the bottom 5 rows of a df

`glimpse()` - gives a brief overview

`view()` - view entire df

`str()` - returns structure of df

`sum(is.na)` - checking for na

length


## 1.2 dplyr functions 

`select()` - selecting variables; can be used with a negative (-x)

`filter()` - filter columns out

`arrange()` - aranges rows

`arrange(desc())` - arranges in descending order

`mutate()` - overwrites / creates a new column - mutate(new_column = old_column + 5)

`group_by()` - creates a group

`summarise()` - creates a summary

can be used with:
`sum()` - summary(x, sum(y))
`count()` - 
`mean()` - 
`n()` - 
`everything()` 

group by and summarise

data <- group_by(column_x, column_y)
summarise(data, column_y = n())

`is.na()` - checks a column for NA values

Useful to use with:
`filter(!is.na(column))` - removes na values from column

`na.drop` - drops NA rows

`pull()` - 

`coalesce()` - 

across() - 

ungroup() - 
tibble() - 
n() - 
na.rm = TRUE; mutate(new_col = old_col = x(na.rm = TRUE)
summarise(count = sum(is.na(col)))



## 1.3 Other important functions

`slice_min()`

`slice_max()`

`across()` 
useful with:
~sum(is.na(.))

`ifelse()` - 
e.g mutate(gender = if_else(gender == "M", "Male", "Female"))
can have a .default

can recode from NA to NA_character
e.g. mutate(compl = recode (compl, "T" = "True", "F" = "False", .default = NA_character_))


as.numeric(name)
as.logical)
as.factor()

## 1.4 Str functions

`str_length()`

`str_extract()`

`str_extract_all()`

`str_detect()`


## 1.5 Joining functions

basic syntax - <type of join>_join(x, y, by = NULL, suffix = c(".x,", ".y"))
```{r}
inner_join(customers, orders, by = "customer_ID")
```

`inner_join()` - only returns records where there's an equivalent in both halves

`left_join()` - all variables from the left df will be included

`right_join()` - opposite of left_join

`full_join()` - includes everything, matching where possible

`semi_join()` - filtering join

`anti_join()` - filtering join

`bind_rows()` - appends rows, e.g. for an updated dataset

`bind_cols()` - appends columns, e.g. for an updated dataset

## 1.5.1 Pivot functions

`pivot_longer()` - takes names_to and values_to

`pivot_wider()` -  takes names_from and values_from

## 1.6 Base R functions

`if_else()`

`sample()`

`all()`

`any()`

`conflicts()`


## 1.6.1 Useful mathematical operators

`%% ` - modulo operator, e.g. 50 %% == 0

`%in%` - checks for if x is in y


## 1.6.2 Conditional operators

`==` - equal to

`!=` - not equal to

`<` - less than

`>` - greater than

`>=` - greater than OR equal to

`%in%` - is in the set


## 1.6.3 Vectorisation of R

The apply family (apply(), lapply(), sapply()) of functions allow you to take a functions and apply it in a vectorised way.

apply() - will try to return the simplest data structure it can (vectors)

lapply() - will usually return a list



## 1.7 Unit testing

Unit testing is creating code to test code that has previously been written.

Tests are generally a good idea for bigger, more advanced functions which are prone to throw more errors.

To unit test in R, apply package testthat.

`expect_equal()` - tests for if values are equal

`expect_error()` - tests for if an error is expect, and will return true if it IS expected.


## 2.1 If statements

Basic syntax
```{r}
if (condition) {
}
```

`if(){}` - if statement (code runs if condition is met)

`if(){if(){}}` - nested if statement

`if(){}else{}` - if else statement

Example
```{r}
if (4 < 5) {
  print("yes it's less")
} else {
  print("no it's not")
}
```

`if(){}else if{}` - else if statement

Example
```{r}
if (x > y) {
  print("yes it's less")
} else if (x < y){
  print("y is greater")
} else {
  print("they're the same")
}
```


## 2.2 For statements

Basic syntax
```{r}
for (num in numbers) {
}
```

`for(){}` - for statement. The () take two statements, a placeholder and an index
`for(){for(){}}` - nested for loop

Example of indexing
```{r}
for (i in 1:length(weekdays)) {
  print(i)
  print("the day is", weekdays[i], sep = " ")
}
```

It is often important to add a counter, e.g.
```{r}
for (i in 1:length(my_vector)) {
  temp_vector[i] <- my_vector[i] + 1
}
```


## 2.3 Which statements
###To be edited ALOT



```{r}
# while loops

while (snowing) {
  clear_drive()
}
```

```{r}
#R example

# set beginning value, and stop value

current_value <- 1
stop_value <- 10

while (current_value < stop_value){
  print(current_value)
  current_val <- current_val + 1
}

```



## 2.4 Functions

```{r}
#functions

#to see code of the functions, run without brackets

library

nchar()

# there are many arguments that can be passed in, but we often skip many of them
# by being explicit, you can move the arguments around, syntax doesn't matter
# what is more common, is passing arguments by place, but when one is skipped, R throws errors



```

```{r}
my_fun <- function(x, y, z) {
      1+1
}

# good practice to not return on the last line
```

```{r}
multiply_by_5 <- function(x) {
  x * 5
}

multiply_by_5(5)
```
```{r}
multiply_by <- function(x, y) {
  x * y
}

multiply_by(1:3, 10)
```

```{r}
beaver1
```


```{r}
make_price <- function(x) {
  paste("£", x)
}
make_price(4)

# paste0 is what you need to remove the space

# round() rounds to two decimal places

```

```{r}
triangle_function <- function(x) {
  (x*(x+1))/2
}

triangle_function(5)
```

```{r}
#global assignment operator


```
```{r}
# to debug functions

#debug() is an option
# but not great because you have to stop it
#undebug()
#debugonce() is better - goes through the code line by line



```




## 3.1 Styling and Best Practice
###To be edited ALOT

Snakecase is best practice
Objects should be nouns and functions should be verbs
Cannot use numbers in objects in the first character, e.g. 1_file
Always start with library calls

Spacing and assignment
Put spaces after commas, and after operations
No space after the start/end of brackets (e.g. (function, bleep) )
No space after calling a function, e.g mean(bleep)
Space before and after assignment operators ( <- )

Line length
Code should not be more than 80 char per line
Code can be split across multiple lines
Should have function name and opening bracket on one line
Args on the next line

Quotation marks
Double and single quotation marks are equally valid as a string
Some strings may have an apostrophe, and then the string breaks
Convention is therefore to use double quotation marks

Comments
Comments are powerful and important in understanding code
Comment above the code


## 3.2 Reproducibility and file paths
###To be edited ALOT

How to Achieve reproducibility

key to be organised, and keep a good folder structure

exam_results[, 1] - This is NOT reproducible 
a more robust/reproducible way is by naming the data, i.e. the column 
exam_results[, "student"]


organising your code so that it is easy to read - one way of doing this
is by creating functions, or modulising it


source function - looks in the working directory for various files
source()
grabs the functon from e.g. an R script, and places in the global env, so it can be run

e.g. 
source("01_multiply_function.R") # this would be a separate script
source("02_add_function.R") # this would be a separate script
a <- multiply_by_n(2, 4)
b <- add_two(a)
b


in an R script, you can add headers / tails
shift + cmd + r
on r script, you can select the little lines menu on top right which automatically indexes



documentation

main points (comments)
purpose and objective - WHY
instruct how to follow the code

key things to note and include:
date (e.g. if packages or similar are or become out of date)
sources
evidence
decision making process
assume the reader has no prior knowledge

restart and run all chunks, is safest way to avoid creating objects etc outside
of the active / required environment/notebook
if you do want to do this, sourcing is a good way to do this
if scripts are very long / take long to run, you can save scripts to a specfic workspace
saveRDS(object, file = "name")
but comment it out after saving it the first time



relative file path

difference between relative file path and absolute file path
relative file path is relative to your working directory, i.e. you do not need
to input the entire filepath


read_csv("/Users/louiseshambrook/GitHub/codeclan_work/codeclan_work/week_02/day_4/w2_d4_reproducability/no_project_example/")
trying to read in data, but unable to locate it / your files have not been set up correctly
by using ~, it brings you to the top of your directory
This is NOT reproducible, as their file structure may differ (and likely will)

if the data file is not accessible on the same level (e.g. if the data is not
contained within the same folder as the rproj, then the library(data) with the 
function data can correct this)
e.g. read_csv(here("data/students")



## 3.3 Assertive programming
###To be edited ALOT

```{r}
# assertive programming
# checking / looking at the data going into your code
```


```{r}
library(tidyverse)
```

```{r}
weather_one <- read_csv("data/delhi_weather_one.csv")
weather_two <- read_csv("data/delhi_weather_two.csv")
```

```{r}
weather_report <- function(weather){
  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

weather_report(weather_one)
```

```{r}
#how will this function handle a different dataset?
weather_report(weather_two)
```

```{r}
library(assertr)

#package to begin to 'fix'/amend our code
```


```{r}
weather_report <- function(weather){
  weather %>%
    verify(temp >= -35 & temp <= 55) %>%
    verify(wind_speed >= 0 & wind_speed <= 108) %>%
    verify(humidity >= 0 & humidity <= 100)

  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

#using verify(), which is similar to filter
```

```{r}
#testing again
weather_report(weather_two)

```

```{r}
#a different function is stop_if_not
```
```{r}
#amending the function again, with stop if not

weather_report <- function(weather){
  stopifnot(
    nrow(weather) > 15,
    ncol(weather) == 4,
    drop_na(weather) == weather
  )
  
  weather %>%
    verify(temp >= -35 & temp <= 55) %>%
    verify(wind_speed >= 0 & wind_speed <= 108) %>%
    verify(humidity >= 0 & humidity <= 100)

  # Our average temperature, humidity and wind speed 
  weather_means <- 
  weather %>% 
  summarise(
    mean_temp = mean(temp),
    mean_humidity = mean(humidity),
    mean_windspeed = mean(wind_speed)
  )

# Show hot and humid days
hot_and_humid <- 
weather %>% 
  filter(temp > 20) %>% 
  filter(humidity > 55)

return(list(weather_means, hot_and_humid))
}

weather_report(weather_one)

```

```{r}
weather_one

```





## 4.1 Tidy data
###To be edited ALOT

```{r}
# What is tidy data?

# Tidy data has structure
# 1. Every variable has its own column
# 2. Each observation has its own row
# 3. Each value must have its own cell (e.g. not Joe Bloggs, 28)

# Tidy data has relationships (e.g. relationship between name and bank account number)

```

```{r}
# Some terminology

# Long format - every row represents an observation belonging to a particular category
# Wide format - observations are spread across multiple columns

```

```{r}
# tidyr package

# pivot_longer() - gathers columns into rows
# pivot_wider() - spreads rows into columns
# seperate() - seperates one column into two
# unite() - unites several into one

```

```{r}
# create subject info tibble
subject_hr <- tibble( name = c("SUBJ01","SUBJ02","SUBJ03"), 
                      hr_am = c(68,72,68), 
                      hr_pm = c(58,89,52))
subject_hr


```

```{r}
subject_hr %>%
  pivot_longer(cols = c("hr_am", "hr_pm"),
               names_to = "measurement_time",
               names_prefix = "hr_",
               values_to = "heartrate")
```


```{r}
income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
                 X10k = c(12,25,3,75), 
                 X10k_to_20k = c(34,14,6,60),
                 X20k_to_30k = c(45,22,10,110))

income

```

```{r}
tidy_income <- income %>%
  pivot_longer(
    -religion,
    names_to = "income_band",
    values_to = "count"
  )
```

```{r}
messy_orders <- tibble( name = c("Joe Bloggs","Jane Smith","Penny Jones"), 
                        order_info = c("XY7282/29-06-19","LO0182/15-03-19","AS6812/04-06-19"), 
                        order_amount = c(58.00,125.00,75.25))

messy_orders

tidy_orders <- messy_orders %>%
  separate(order_info, 
           c("order_number", "date"),
           sep = "/"
  )

```

```{r}
hospital_visits <- read_csv("3_tidy_data/data/hospitals93to98.csv")
```

```{r}
hospital_visits_long <- hospital_visits %>%
  pivot_longer(cols = starts_with("FY"),
               names_to = "year",
               values_to = "visit_numbers",
               names_prefix = "FY"
  )
```

```{r}
hospital_visits_wide <- hospital_visits_long %>%
  pivot_wider(names_from = "Field",
              values_from = "visit_numbers"
  )

```

```{r}
hospital_visits_coded <- hospital_visits_wide %>%
  separate(IcdChapter,
           c("code", "description"), sep = "\\.")

```


```{r}
hospital_visits_unite <- hospital_visits_coded %>%
  unite(disease,
        c("code", "description"),
        sep = ":"
  )

```



## 5.1 Regular expressions (regex)
###To be edited ALOT

```{r}
# Regex - regular expressions

library(tidyverse)
```

```{r}
single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")

```

```{r}
pattern <- "a"
```

```{r}
single_string %>%
  str_detect(pattern)

strings %>%
  str_detect(pattern)
```

```{r}
pattern <- "str"
strings %>%
  str_detect(pattern)
```
```{r}
pattern <- "[abc]"
strings %>%
  str_detect(pattern)

# the square brackets are what make it regex: it's looking for a or b or c
```

```{r}
pattern <- "[a]" # look for letters between a and c
strings %>%
  str_detect(pattern)
```

```{r}
pattern <- "[a-zA-Z]" # look for lower and upper case a-z (if you included ",", it would look for commas)
strings %>%
  str_detect(pattern)
```

```{r}
pattern <- "i{3}" # find where there's 3 consecutive lowercase i's
strings %>%
  str_detect(pattern)
```

```{r}
# Task - 10 mins
# Make a vector containing the following strings:
# “We’re having great fun learning about data analysis at CodeClan!”
# “CodeClan is at 37 Castle Terrace in Edinburgh, ClockWise in Glasgow, and Wasps Academy in Inverness”
# “Our cohort code starts with DE, DG or D”
# “You get a great view of the castle or the national conservatoire from just outside CodeClan’s door”
# Write and test regular expressions which will match:
# Any string mentioning CodeClan
# Any string containing two consecutive capital letters
# Any string not including the word “castle” or “Castle”
```

```{r}

test_strings <- c(
  "We're having great fun learning about data analysis at CodeClan!",
  "CodeClan is at 37 Castle Terrace in Edinburgh, ClockWise in Glasgow, and Wasps Academy in Inverness",
  "Our cohort code starts with DE, DG or D",
  "You get a great view of the castle or the national conservatoire from just outside CodeClan's door"
)

```

```{r}

test_pattern <- "CodeClan"
test_strings %>%
  str_detect(test_pattern)

test_pattern_two <- "[A-Z]{2}"
test_strings %>%
  str_detect(test_pattern_two)

!str_detect(test_strings, "[cC]astle")

```

```{r}
# recap

# [] - or (in strings)
# {} - number of occurrences

```

```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```

```{r}
# (?i) negates the capitalizaton

# "(?i)[a-zA-Z]"
```

```{r}
pattern <- "[a-z]"
strings %>%
  str_extract(pattern)

# is vectorised, i.e it was applied to each vector 
```

```{r}
pattern <- "[a-z]@"
strings %>%
  str_extract(pattern)
```

```{r}
# further modifiers

# * - zero or more occurences
# + - one or more occurences
```

```{r}
invalid_email <- "@emailprovider.com"
pattern <- "[a-z]*@emailprovider.com"

invalid_email %>%
  str_extract(pattern)

#ALSO includes none, basically a default behaviour
```

```{r}
invalid_email <- "@emailprovider.com"
pattern <- "[a-z]+@emailprovider.com"

invalid_email %>%
  str_extract(pattern)

#will fail becuase there's nothing before the @
```

```{r}
pattern <- "[a-z]+@"
strings %>%
  str_extract(pattern)
```

```{r}
pattern <- "[a-z]+@"
strings %>%
  str_extract(pattern)

pattern <- "[a-z0-9]+@[a-z]+"
strings %>%
  str_extract(pattern)

```

```{r}
pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"
strings %>%
  str_extract(pattern)

```

to get all the email addresses, we need to run a different function
 
strings %>%   str_extract_all(pattern)


```{r}
#changing email addresses

pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"
strings %>%
  str_replace(pattern, "REDACTED")


pattern <- "[a-z0-9_\\.]+@[a-z\\.]+"
strings %>%
  str_replace_all(pattern, "REDACTED")

```

```{r}
#task
# Make a vector containing the strings:
# “Colin’s mobile number definitely isn’t 07777777777.”
# “A landline might be 01010 123456.”
# “Mhairi, Del, Aileen and Steph have been some of your instructors so far.”
# “I’m called your_name_here and I live in your_town_here.”
# Use regular expressions to:
# Extract all the names
# Extract the landline number
# Insert your name and where you live into the fourth string

task_string <- c("Colin’s mobile number definitely isn’t 07777777777.”,
                  “A landline might be 01010 123456.”,
                  “Mhairi, Del, Aileen and Steph have been some of your instructors so far.”,
                  “I’m called your_name_here and I live in your_town_here.")

# Extract all the names
name_pattern <- "[A-Z]+[a-z,]+[ ]"
task_string %>%
  str_extract_all(name_pattern)

number_pattern <- "[0-9]{5}"
task_string %>%
  str_extract_all(number_pattern)

# ANSWERS
name_pattern <- "[A-Z][a-z]+"
task_string %>%
  str_extract_all(name_pattern)

#pattern is that;
#it starts with a capital, followed by a lowercase, and there is more than one occurrence

#ANSWER FOR NUMBER TWO
number_pattern <- "[0-9]{5} [0-9]{6}"
task_string %>%
  str_extract_all(number_pattern)

#ANSWER FOR WHAT I WAS TRYING TO DO
number_pattern <- "[0-9]{5} [0-9]{6}|[0-9]{11}"
task_string %>%
  str_extract_all(number_pattern)

my_address <- "Edinburgh"
my_name <- "Louise"
task_string %>%
  str_replace("your_name_here", my_name) %>%
  str_replace("your_town_here", my_address)

```

```{r}
punct_strings <- c(
  "Why is this a question?", 
  "I am a sentence with no punctation",
  "I am a well punctuated sentence.", 
  "WOW!"
)
```

```{r}
#punctuation strings

punct_strings %>%
  str_extract_all("[:punct:]")

punct_strings %>%
  str_extract_all("[A-Z]{3}[:punct:]") %>%
  flatten_chr()

```

```{r}
# using the ^ - denotes the start of the string
# and $ - denotes the end of the string
```

```{r}
test_strings <- c(
  "abc", 
  "Abc",
  "abC",
  "s"
)

test_strings %>%
  str_extract_all("[a-z]")
```

```{r}
# if we only want where the string starts with x 

test_strings %>%
  str_extract_all("^[a-z]")
```

```{r}
# if we want it to end with x

test_strings %>%
  str_extract_all("[a-z]$")

```

```{r}
# if we want a string in all xx (e.g. lowercase, so start END end with)
# only returns one ("s"), because we don't use the +

test_strings %>%
  str_extract_all("^[a-z]$")
```

Dealing with strings within a tibble

```{r}
student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")

student_info <- tibble(student, age, emails)

student_info 
```

```{r}
student_info %>%
  mutate(email_detect = str_detect(emails, "[a-z0-9]+@"))
```

```{r}
student_info %>%
  mutate(emails = str_extract(emails, "[a-z0-9_\\.]+@[a-z\\.]+"))
```



```{r}
# Question 1.
# Are the Thought Police ever mentioned?

whole_text %>%
  str_extract_all(
    "(?i)thought police"
  )

# ANSWER
# pattern <- "[tT]hought +[pP]olice"
# 
# whole_text %>%
#   str_detect(pattern)
```

```{r}
# Question 2.
# Count how often Big Brother is mentioned

whole_text %>%
  str_extract_all("(?i)big brother") %>%
  flatten_chr()

# becomes a vector (therefore need to count length / number of instances in the vector)

#20

# ANSWER
# pattern <- "[Bb][Ii][Gg] +[Bb][Rr][Oo][Tt][Hh][Ee][Rr]"
# 
# whole_text %>%
#   str_extract_all(pattern) %>%
#   flatten_chr() %>%
#   length()
```

```{r}
# Question 3.
# The chapter is written from the point of view of the protagonist, Winston. Does he ever mention the other main character (Julia) by name?

whole_text %>%
  str_extract_all("(?i)Julia")

OR

whole_text %>%
  str_detect("(?i)Julia")

# ANSWER
# pattern <- "[jJ]ulia"
# 
# whole_text %>%
#   str_detect(pattern)

```

```{r}
# Question 4.
# Winston writes the date in a specific format: month nth, year eg. August 5th, 2019. Write a regular expression to match dates in that format and find the date written in his diary.

whole_text %>%
  str_extract_all("[A-Za-z]+[ ][0-9a-z]+,[ ][0-9]+")

# ANSWER
# pattern <- "[A-Z][a-z]+ [0-9]+(th|rd|st|nd), [0-9]{4}"
# whole_text %>%
#   str_extract(pattern)
```

```{r}
# Question 5.
# Winston wants some anonymity, redact his name somehow.

whole_text %>%
  str_extract_all("(?i)Winston") %>%
  flatten_chr() %>%
  str_replace_all("Winston", "REDACTED")


# ANSWER
# pattern <- "[wW]inston"

# whole_text <- whole_text %>%
#   str_replace_all(pattern, "REDACTED")

# #check that there are no winstons mentioned
# whole_text %>%
#   str_detect(pattern)

```

```{r}
# Question 6.
# Stocks of Victory Gin have run out and the people of Oceania will have to drink vodka instead. Replace “gin” with “vodka”, matching case wherever possible. There’s still loads of Victory Coffee, though, so don’t replace that.

#my answer/suggestion - did not give the correct words
# whole_text %>%
#   str_extract_all("gin") %>%
#   flatten_chr() %>%
#   str_replace_all("gin", "vodka")

pattern <- "[ ,\\.-][gG][iI][nN][ ,\\.-]"
whole_text %>%
  str_extract_all(pattern)

whole_text %>%
  str_extract_all("Vodka")


# ANSWER
#pattern <- "[ ,\\.-][gG][iI][nN][ ,\\.-]"
# whole_text %>%
#   str_extract_all(pattern) %>%
#   flatten_chr()
# 
# whole_text <- whole_text %>%
#   str_replace_all(" GIN.", " VODKA.") %>%
#   str_replace_all(" gin.", " vodka.") %>%
#   str_replace_all(" gin ", " vodka ")
```

```{r}
#Q7 - did not get this far

whole_text <- whole_text %>%
  str_replace_all("Department", "Section") %>%
  str_replace_all("Ministry", "Department") %>%
  str_replace_all("Ministrie", "Department") %>%
  str_replace_all("Mini", "Dep")
```





## 5.2 Strings
###To be edited ALOT

```{r}
# defining strings

string <- 'this is a string'
#throws an error if you use an apostrophe

double_quotes <- "this is also a string"

```

```{r}
library(tidyverse)
```

```{r}
employees <- tibble(
  first_name = c("Sally", "steven", "Mary", "John"),
  last_name = c("Smith", "hasTings", "Milne", "Davidson"),
  postcode = c("EH10 6GH", "EH3 7HL", "G1 7FR", NA), 
  location = c(" ", "Edinburgh", "", "Stirling")
)

employees

# Sally and Mary both don't have a location, but Sally does technically have a
# location, which is a " " 
# so ALWAYS check for empty spaces

```

```{r}
# checking for empty spaces

employees %>%
  group_by(location) %>%
  summarise(count = n())

# better would be to go through and change to NA

```

```{r}
# Manipulating strings

str_c("this", "is", "a", "string")
# concatenates the input

```

```{r}
# function can be used to create a name (instead of first_name and last_name)

employees %>%
  mutate(full_name = str_c(first_name, last_name, sep = " "))
```

```{r}
# fixing casing / casing
# janitor / clean_names for column is one option

pangram <- "the quick brown fox jumps over the lazy dog"

#all in uppercase
pangram %>%
  str_to_upper()

#all in lowercase
pangram %>%
  str_to_lower()

#all words start with uppercase
pangram %>%
  str_to_title()

# letters after full stop begin with capital
pangram %>%
  str_to_sentence()

```

```{r}
#fixing the case of names

employees <- employees %>%
  mutate(first_name = str_to_title(first_name),
         last_name = str_to_title(last_name),
         full_name = str_c(first_name, last_name, sep = " ")
  )

```

```{r}
unbroken_string <- "we're going to split this string up"

unbroken_string %>%
  str_split(" ") %>%
  flatten_chr()


```

```{r}
employees %>%
  mutate(postcode_new = str_split(postcode, " ")
  ) %>%
  unnest(postcode_new)

# not very useful, but the separate function is more useful
```

```{r}
employees %>%
  separate(postcode, c("postcode_outer", "postcode_inner"),
           sep = " ")
```

```{r}
# Task - 10 mins
# Concatenate the strings "Manipulating" "strings" "is" "fun!"
# Capitalise the result and save it in a new variable
# Split the formatted string into individual characters

string_task <- str_c("Manipulating", "strings", "is", "fun!")
# can add sep = " "

capitalised_string_task <- str_to_upper(string_task)

str_split(capitalised_string_task, "") %>%
  flatten_chr()

```

```{r}
# finding strings / substrings

full_string <- "let's grab a substring from things"

full_string %>%
  str_sub(start = 1, end = 5)

# counts from the start of the index of the string, to the end
# 1 index language (not 0)
# doesn't require an end argument, will count towards the end
# end of the string is -1

full_string %>%
  str_sub(-5, -1)

```

```{r}
employees %>%
  mutate(postcode_inner = str_sub(postcode, -3))

# starts 3 points from the end

employees %>%
  mutate(postcode_outer = str_sub(postcode, start = 1, end = -4))
```

```{r}
#replace bits of string

original_string <- "we're going to replace some words"

# string_sub() is an option, but because it's hardcoding, it isn't great

str_sub(original_string, 1, 22) <- "I've changed"

```

```{r}
#numbers as strings

number <- 123456.789

number %>%
  format(nsmall = 2)

number %>%
  format(nsmall = 2, big.mark = ",", decimal.mark = ".")

#changed from numeric, to string, with 2 decimal places
# can be used for graphing, for the sake of formatting
```

```{r}
# Here are some more questions for you to practice with. Remember, the point of these questions is not to do something that makes sense for this dataset, but rather to allow you to practice using the techniques we’ve learnt today on a real dataset. When it comes to your projects, you will be in charge of deciding how to work with the strings in your dataset.
# We are going to use some Spotify data for this task:
# Uppercase the text in the playlist_name column. Overwrite the existing column to store the result.
# Lowcase the text in the track_artist column. Store the result in a new column called track_artist_lowcase.
# Take the first 5 characters of the playlist_id column, and save it in a new column called playlist_start.
# Take the first 4 characters of the track_album_release_date column and save it in a new column called year.
# What else can you do to this dataset to practice? Can you work with the numbers? This is open ended - have fun!

```

```{r}
spotify_songs <- read_csv("data/spotify_sample.csv")
```
```{r}

spotify_cleaned <- spotify_songs %>%
  clean_names() %>%
  mutate(playlist_name = str_to_upper(playlist_name)) %>%
  mutate(track_artist_lowcase = str_to_lower(track_artist)) %>%
  mutate(playlist_start = str_sub(playlist_id, start = 1, end = 5)) %>%
  mutate(year = str_sub(track_album_release_date, start = 1, end = 4))

```


## 6.1 Matricies and lists
###To be edited ALOT


```{r}
# Matrices and lists

# Data structures

# vectors
num <- 1:5
strings <- c("This is a string", "hello")
boolean <- c(TRUE, FALSE)

# atomic vector - hold one type of data
# versus list vector, aka list

class(nums)
```
```{r}
list("text", TRUE, c(1, 2, 3), list("meep", 4))
```

```{r}
# as.list() - is the other way of making lists
```

```{r}
# indexing and accessing a list - subsetting lists

x <- list(1, "a", TRUE, c(1, 2, 3))

x[[2]]

# [] will always return the same data structure that you were trying to access
```

```{r}
# can also be accessed by naming lists

# $ denotes that a is a name in the list

named_list <- list(a = "I love R", b = 1:10)

named_list$a
```

```{r}
# access through []

named_list[["a"]]
```

```{r}
# Nested lists

# you can lists within a list inside another list

student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info

# extract the data/lists

student_info$Bobby$height
#gives bobby's height

student_info[["Agnes"]][["shoe_size"]]
```

```{r}
#Task

# Task - 5 minutes
# 
# Make a named list that contains a numeric vector and a character vector.
# 
# Access the second element of your list using double square brackets [[]]. What is it's class?
# 
# Access the second element of your list using single square brackets []. What is it's class?
# 
# Access the second element of your list by name. What is it's class?
```

```{r}
task_list <- list(a = c(4, 5, 6), b = c("meep", "hello", "here"))

task_list[[2]]

class(task_list[[2]])
  
task_list[2]

class(task_list[2])

task_list$b
class(task_list$b)

```

```{r}
# Matrices

# a data structure that holds rectangular data

matrix(1:4, nrow = 2, ncol = 2)

# if we want it populated rowwise (default is by column)

matrix(1:4, nrow = 2, ncol = 2, byrow = TRUE)
```

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)

class(m)

# key distinction between matrices and DF's - matrices they MUST be the same type
```

```{r}
# selecting elemnts from the matrix
m[1,3]

# select entire row
m[1,]

#select entire column
m[,3]


```
```{r}
# you can select multiple rows or columns by passing a vector

m[, c(1,3)]
#returns all rows for 1st and 3rd column
```

```{r}
# advantage of matrix, is to be able to perform calculations on entire matrix

m * 5
```

```{r}
library(dplyr)
library(magrittr)

numers <- iris %>% select(-Species)

# if it's all the same data type
numers * 3

# access the first row of a df
numers[1, ]
```

```{r}
# Working with rows and columns

# want to get the mean average of a column

#colMeans()

colMeans(m)
```

```{r}
# Other convenience functions

`rowSums`

`colSums`

`rowMeans`

rowSums(m)
colSums(m)

```
```{r}
# Task
# Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
# Find the element in the 2nd row, 3rd column.
# Check that all the columns are identical by checking that their sums are the same.
```

```{r}
#Make a matrix

my_matrix <- matrix(1:3, nrow = 3, ncol = 4)

my_matrix[2,3]

colSums(my_matrix)


```



## 10.1 Code chunks from homework

## 10.2 Weekend homework 1
###To be edited ALOT

```{r}
# Creating my main dataframe - removing row_ID, ISBN, ISBN13,
# publication_date and text_reviews_count
books <- read_csv("data/books.csv")
main_books_data <- books %>%
  select(-rowid, -isbn, -isbn13, -text_reviews_count, -bookID)

#Q1 - What is the highest rated book from the original series in English by JK?
highest_hp_rating <- main_books_data %>%
  select(title, authors, average_rating) %>%
  filter(str_detect(authors, "Rowling"),
         str_detect(title, "Harry Potter and the"),
         !str_detect(title, "Sorcerer's")) %>%
  group_by(average_rating, title) %>%
  distinct(title) %>%
  ungroup() %>%
  slice_max(1)

#Q2 - Is the average of HP1-7 higher than the average of LotR?
# Calculating the Harry Potter average
hp_avg_rating <- main_books_data %>%
  select(title, authors, average_rating) %>%
  filter(str_detect(authors, "Rowling"),
         str_detect(title, "Harry Potter and the"),
         !str_detect(title, "Sorcerer's")) %>%
  distinct(title, average_rating) %>%
  summarise(avg_rating = mean(average_rating))
hp_avg_rating

# Calculating the LotR average
lotr_avg_rating <- main_books_data %>%
  select(title, authors, average_rating, publisher) %>%
  filter(str_detect(title, "The Return of the King|The Fellowship of the Ring|The Two Towers"),
         str_detect(publisher, "Houghton Mifflin|Mariner Books"), 
         authors != "Jude Fisher/Viggo Mortensen",
         publisher != "Houghton Mifflin Harcourt"
  ) %>%
  summarise(avg_rating = mean(average_rating))
lotr_avg_rating

higher_hp_lotr_rating <- hp_avg_rating > lotr_avg_rating
higher_hp_lotr_rating
#TRUE - HP is higher than LOTR

#Q3 What are the ratings and titles of the top 5 rated french books
#ranked by number of ratings (not rating)
top_french_books <- main_books_data %>%
  select(title, average_rating, ratings_count, language_code) %>%
  filter(language_code == "fre") %>%
  group_by(ratings_count, average_rating) %>%
  arrange(desc(ratings_count))

head(top_french_books, 5)

#top_french_books gives the output

#Q4 - What is the mean rating for books published by Puffin?
mean_puffin_rating <- main_books_data %>%
  select(title, average_rating, publisher) %>%
  summarise(puffin_avg_rating = (mean(average_rating)))

# answer is 3.934075

#Q5 - Are shorter books, rated lower than longer books?
# low book rating 
short_book <- main_books_data %>%
  select(title, average_rating, num_pages) %>%
  mutate(length_of_book = case_when(
    num_pages < 200 ~ "under 200",
    num_pages > 200 & num_pages < 400 ~ "200 to 400",
    TRUE ~ "above 2000")) %>%
  filter(length_of_book == "200 to 400") %>%
  summarise(low_book_avg_rating = mean(average_rating))

long_book <- main_books_data %>%
  select(title, average_rating, num_pages) %>%
  mutate(length_of_book = case_when(
    num_pages < 600 ~ "under 200",
    num_pages > 600 & num_pages < 800 ~ "600 to 800",
    TRUE ~ "above 2000")) %>%
  filter(length_of_book == "600 to 800") %>%
  summarise(high_book_avg_rating = mean(average_rating))

highest_book_split_rating <- short_book < long_book
highest_book_split_rating
#TRUE

#OR
book_summ <- books %>%
  select(title, average_rating, num_pages) %>%
  mutate(length_of_book = case_when(
    num_pages < 200 ~ "0. under 200",
    num_pages >= 200 & num_pages < 400 ~ "1. 200 to 400",
    num_pages >= 400 & num_pages < 600 ~ "2. 400 to 600",
    num_pages >= 600 & num_pages < 800 ~ "3. 600 to 800",
    TRUE ~ "4. above 800")) %>%
  # filter(length_of_book == "200 to 400") %>%
  group_by(length_of_book) %>%
  summarise(avg_rating = mean(average_rating))
```

### NOTES TO TIDY
library(MASS)
select(mtcars, mpg) - will give an error if MASS is loaded
error is that it's trying to run the select function from MASS
dplyr::select(mtcars, mpg) - to load a function directly from the package

separate()
unite()
rowSums(m)
colSums(m)
sapply()
lapply()
apply()

purrr library

rename()