---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

```{r}
scotland_level <- read_csv("data/opendata_inc9519_scotland.csv")
by_region <- read_csv("data/opendata_inc9519_region.csv")
by_health_board <- read_csv("data/opendata_inc9519_hb.csv")
label_data <- read_csv("data/geography_codes_and_labels_hb2014_01042019.csv")

by_health_board <- clean_names(by_health_board)

```


In order to help inform the planning for provision of cancer treatment services in NHS Borders, we would like to gain better understanding of the incidence of cancer in NHS Borders.

To begin to answer / explore this 
 - I assumed that the data was new, i.e. that this would only include new diagnoses / 
 stats from this year
 - This data goes back to 1995 

## Question 1 - What are the 5 most common / prevalent types ?

```{r}
by_health_board %>%
  select(1:7) %>%
  filter(hb == "S08000016", cancer_site != "All cancer types") %>%
  group_by(cancer_site) %>%
  summarise(sum_type = sum(incidences_all_ages)) %>%
  arrange(desc(sum_type)) %>%
  head(5) %>%
  ggplot(aes(x = cancer_site, y = sum_type)) +
  geom_col() +
  coord_flip()

```
Because there are a large amount of types, to present the information in a reasonable limit, it was necessary to set a limit, so this was set at 5.
These numbers are calculated across 1995-2019, to gain a basic insight into whether there is a prevalent type, and if so, which. 


## Question 2 - How does this break down across time?

```{r}
by_health_board %>%
  select(1:7) %>%
  filter(hb == "S08000016", cancer_site != "All cancer types",
         cancer_site == "Breast" | 
         cancer_site == "Trachea, bronchus and lung" |
          cancer_site == "Non-melanoma skin cancer" |
          cancer_site == "Colorectal cancer" | 
          cancer_site == "Basal cell carcinoma of the skin",
         sex == "All"
         ) %>%
  ggplot(aes(x = year, y = incidences_all_ages, fill = cancer_site)) +
  geom_col() +
  facet_wrap(~cancer_site)
  
  
```

## Question 3 - How does this break down across demographics? (gender)

```{r}
by_health_board %>%
  select(1:7) %>%
  filter(hb == "S08000016", cancer_site != "All cancer types",
         cancer_site == "Breast" | 
         cancer_site == "Trachea, bronchus and lung" |
          cancer_site == "Non-melanoma skin cancer" |
          cancer_site == "Colorectal cancer" | 
          cancer_site == "Basal cell carcinoma of the skin",
         sex == "Female" | sex == "Male"
         ) %>%
  ggplot(aes(x = year, y = incidences_all_ages, fill = sex)) +
  geom_col()
```

## Question 4 - Statistically - are females or males diagnosed with cancer in the NHS Borders region more?
(in this section that I looked at)

```{r}

```

