---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

```{r}
scotland_level <- read_csv("data/opendata_inc9519_scotland.csv")
by_region <- read_csv("data/opendata_inc9519_region.csv")
by_health_board <- read_csv("data/opendata_inc9519_hb.csv")
label_data <- read_csv("data/geography_codes_and_labels_hb2014_01042019.csv")



```


In order to help inform the planning for provision of cancer treatment services in NHS Borders, we would like to gain better understanding of the incidence of cancer in NHS Borders.

To begin to answer / explore this 
 - I assumed that the data was new, i.e. that this would only include new diagnoses / 
 stats from this year
 - This data goes back to 1995 

Children are my interest, so that will lead my focus.
Divide into two groups; children, and young people (this is how PHS divides them up).

## Question 1 - What is the most prevalent type of cancer in each group?
```{r}
# by_region - filter by south east 
# child group



# by_region - filter by south east
# young people group
```

## Question 2 - How are the demographics of each group?

```{r}
# by_region - how are these demographics split?
# child group



# by_region - how are these demographics split?
# young people group

```

## Question 3- Has the prevalence changed between 1995 and 2019?

```{r}
#by_region - child group - most prevalent 1995 / 2019



#by_region - young people group - most prevalent 1995 / 2019

```

## Question 4 - 

```{r}

```

