---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


Week7 day 1

Dates and Times

Why are they difficult?
- formatting them in programming can be tricky
-lots of different ways to enter them
- tricky data format to work with


ISO-8601 Format
- YEAR-MONTH-DAY

Format: YEAR-MONTH-DAY-HH-MM-SS
Min digits: 4 b/c of 2000
Clock system: 24 hours
Time formats: 4 
Timezone - UTC
Combined date-times - 2021-12-13-T10:13

```{r}
library(lubridate)
```

```{r}
my_bday <- "09-01-2021"
my_bday
```

```{r}
date <- dmy(my_bday)
# transforms the string to a date
```

```{r}
mdy("January 9th, 2021")
```

```{r}
mdy_hms("01-09-21 10:18:26")
```

```{r}
ymd(20190109)
```

```{r}
netflix <- read_csv("data/netflix.csv")
```

```{r}
netflix <- netflix %>%
  mutate(date_added = mdy(date_added))
```

```{r}
# extracting exact days / months from the date
netflix <- netflix %>%
  mutate(
    day = day(date_added),
    month = month(date_added, label = TRUE, abbr = FALSE),
    year = year(date_added),
    weekday = wday(date_added, label = TRUE, abbr = FALSE)
  )

# weekdays() and wday() is the same - weekdays() is base R and wday() is lubridate
```

```{r}
netflix <- netflix %>%
  mutate(made_date = make_datetime(year, month, day))

head(netflix)
```

Add a time
```{r}
netflix <- netflix %>%
  mutate(made_date = make_datetime(year, month, day, hour = 16, min = 59, sec = 45))
```

Calculating time spans

- difftime: represents 
- period
- durations

```{r}
netflix <- netflix %>%
  mutate(time_difftime = today() - date_added) %>%
  select(time_difftime, date_added)
```

Period:
Subsetting / breaking down periods into readable / digestable periods

```{r}
netflix <- netflix %>%
  mutate(time_period = as.period(today() - date_added),
         time_years = as.numeric(time_period, "years"),
         time_hours = as.numeric(time_period, "hours"),
         time_mins = as.numeric(time_period, "minutes"),
         time_seconds = as.numeric(time_period, "seconds"),
         )
```
When creating/reporting periods, it is important to be careful/cautious about
leap years. Because leap years don't just appear out of nowhere, it is important
to be aware of small changes, i.e seconds/minutes/hours, etc - each year
is 365.25 years 
It is really only important to watch that you get the right amount of days


Netflix promotes new releases for 2 weeks before release
Netflix takes shows/movies down after 4 months
```{r}
netflix <- netflix %>%
  mutate(promo_date = date_added - weeks(2),
         removal_date = date_added + months(4))

netflix %>%
  select(date_added, promo_date, removal_date)
```


Durations:
Records the exact passage of time since a date
```{r}
netflix <- netflix %>%
  mutate(time_duration = as.duration(today() - date_added))

netflix %>%
  select(date_added, time_difftime, time_years, time_duration)
```


Timezones
```{r}
ymd_hms("2021-12-31 11:30:22", tz = "Atlantic/Faroe")
```


Force_tz - add a timezone, doesn't change date/time
with_tz - figures out what date and time would be if you were in that time zone
```{r}
netflix <- netflix %>%
  mutate(date_tz1 = force_tz(date_added, tzone = "Asia/Singapore"),
         date_tz2 = with_tz(date_added, tzone = "Asia/Singapore"))

netflix %>%
  select(date_added, date_tz1, date_tz2)
```


Lesson 2

Timeseries

Part of the PDA

- continuous data over time
- time is the primary axis
- append rather than replace


```{r}
library(tsibble)
library(tsibbledata)
library(feasts)
library(fable)
```

tsibble is for working with timeseries data
feasts - features etc for statistics

index_var - find the index of a tsibble
key_vars - find the key of a tsibble

You can use dplyr verbs with tsibble, but you cannot deselect / drop time as a 
column

```{r}
# extra functions to filter dates
vic_year <- vic_elec %>%
   filter_index("2012-01")

```

```{r}
# vic_year %>%
#   group_by(year) %>%
#   summarise(mean_temp = mean(Temperature))

vic_year %>%
  index_by(Time) %>%
  summarise(mean_temp = mean(Temperature))

you need to index instead of group
```

Make plots
```{r}
elec_date <- vic_elec %>%
  index_by(Date) %>%
  summarise(mean_temp = mean(Temperature))

elec_date %>%
  ggplot() +
  aes(x = Date, y = mean_temp) +
  geom_line()
```

```{r}
elec_month <- vic_elec %>%
  mutate(month = month(Date, label = T)) %>% 
  index_by(month) %>% 
  summarise(mean_temp = mean(Temperature)) %>% 
  ggplot(aes(x = month, y = mean_temp)) +
  geom_line(group = 1)
```

```{r}
library(slider)
```

```{r}
# calculate a rolling window
# the . in .f corresponds to the current window in the data, i.e. each subset
elec_rolling <- vic_elec %>% 
  mutate(
    temp_moving_avg = slide_dbl(
      .x = Temperature, 
      .f = ~ mean(., na.rm = TRUE),
      .before = 2000,
      .after = 2000,
      .complete = TRUE
    )
  )

elec_rolling
```

```{r}
# plot the data 
ggplot(elec_rolling) + 
  geom_line(aes(x = Date, y = Temperature), colour = "grey") + 
  geom_line(aes(x = Date, y = temp_moving_avg), colour = "red")
```

The slide_dbl() function makes a number of arguments available, so let’s play around with them a little!
Try setting .before and .after to 100. How does the plot change?
What happens if you set .complete = TRUE?


Feasts package
- useful for plotting


```{r}
index_var(tourism)
key_vars(tourism)
```


```{r}
holidays <- tourism %>%
  filter(Purpose %in% c("Holiday","Business")) %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))

holidays
```

```{r}
holidays %>%
  autoplot(Trips)
```

```{r}
holidays %>%
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>%
  gg_season(Trips)
```


```{r}
holidays %>%
  filter(State %in% c("Queensland", "New South Wales", "Victoria")) %>%
  gg_subseries(Trips)
```


