---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)

total_sales <- read_csv("total_sales.csv")
got_ratings <- read_csv("got_ratings.csv")
table_of_numbers <- read_csv("table_of_numbers.csv")
monthly_sales <- read_csv("monthly_sales.csv")
```

Day 2
Making effective plots

Cleveland Mcgill scale

Remaking the plot from the notes, but in a bar plot
```{r}
ggplot(total_sales) +
  aes(x = branch, y = sales) +
  geom_col() +
  coord_flip()
```

changing the order of the bars with reorder()
```{r}
ggplot(total_sales) +
  aes(x = reorder(branch, sales), y = sales) +
  geom_col() +
  coord_flip()
```

Preattentive attributes
- Size
- Colour
- Orientation
- Added marks (underlining)
- Enclosure

```{r}
ggplot(total_sales) +
  aes(x = reorder(branch, sales), y = sales, fill = branch == "Edinburgh") +
  geom_col() +
  scale_fill_manual(values = c("grey60", "light blue"), guide = "none") +
  coord_flip() 

```

Annotating graphs
```{r}
total_sales_annotated <- total_sales %>%
  mutate(label_text = ifelse(branch == "Edinburgh", "Edinburgh branch has 620k sales", NA))

ggplot(total_sales_annotated) +
  aes(x = reorder(branch, sales), y = sales, fill = branch == "Edinburgh") +
  geom_col() +
  scale_fill_manual(values = c("grey60", "light blue"), guide = "none") +
  coord_flip() +
  geom_text(aes(label = label_text), color = "darkblue", size = 4, hjust = -0.1)

```

TASK
Rewrite the graph above in a more compelling manner. Consider how you could emphasise which branches have been most successful in the month of April compared to January.
```{r}

monthly_sales %>%
  mutate(month = factor(month, levels = c("Jan", "Feb", "March", "April"))) %>%
head(monthly_sales)

monthly_annotated <- monthly_sales %>%
  mutate(label = ifelse(branch == "Leeds", "Leeds improved", NA),
         label2 = ifelse(branch == "Glasgow", "Glasgows improved", NA),
         label3 = ifelse(branch == "Manchester", "Edinburgh improved", NA))

ggplot(monthly_sales) +
      aes(x = reorder(branch, sales), y = sales, fill = month) +
      geom_col(position = "dodge") +
      coord_flip() +
      scale_fill_manual(values = c("lightgrey", "skyblue", "lightgrey", "blue"), guide = "none") +
      geom_text(aes(label1 = label), color = "darkblue", size = 4, hjust = -0.1) +
      geom_text(aes(label2 = label), color = "darkblue", size = 4, hjust = -0.1) +
      geom_text(aes(label3 = label), color = "darkblue", size = 4, hjust = -0.1)

```

TASK
```{r}
ggplot(monthly_sales) +
  aes(x = month, y = sales, group = branch, colour = branch) +
  geom_line() 
```
Questions to think about
Which graph is superior for comparing sales growth between branches?  
Which graph is superior for comparing sales in April between branches?
A colleague proposes using the grouped bar chart but changing the order of branches between months so they are ordered by sales numbers in that month. Comment on this suggestion.
A senior manager suspects the proportion of sales arising from Northern offices (Edinburgh, Glasgow, Leeds, Manchester) are growing relative to southern offices. Produce a visualisation that speaks to this.


