---
title: "R Notebook"
output: html_notebook
---

heart lab

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

```{r}
heart_data <- read_csv("heart.csv")
```

1. Load and explore the data held in heart.csv. A data dictionary is also available.
```{r}
glimpse(heart_data)
```

2. Plot a histogram and/or boxplot for thalach (maximum heart rate achieved in stress testing). How normal is the distribution?
```{r}
heart_data %>%
  ggplot(aes(x = thalach)) +
  geom_histogram()

# q answer - pretty normal - but slightly left skewed - kinda tapers off.
```
3. Also plot a barplot of exang (whether the patient displayed exercise induced angina)
```{r}
heart_data %>%
  ggplot() +
  geom_bar(aes(x = exang))

# this is very uneven - 
```

4. Now calculate population parameters:
what is the mean thalach in the population.
What proportion of patients have exercise induced angina (exang)?
```{r}
summary_pop_param <- heart_data %>%
  summarise(mean_thalach = mean(thalach),
            prop_exang = sum(exang == 1) / n())

```
5. Now take a single sample of 50-patients from the population and calculate point estimates for mean thalach and proportion of exang from this sample (let’s call these mean_thalach and prop_exang going forward).
[Hint - use rep_sample_n() from the infer package to do this]
```{r}
# sample
heart_sample <- heart_data %>%
  rep_sample_n(size = 50, reps = 1)

#mean thalach and exang

point_estimate_sample <- heart_sample %>%
  ungroup() %>%
  summarise(mean_thalach = mean(thalach),
            prop_exang = sum(exang == 1) / n())

```

6. What are the sampling errors of these point estimates?
[Hint - remember a sampling error is defined as the difference between a point estimate drawn from a sample, and a population parameter]
```{r}
summary_pop_param - point_estimate_sample

```
7. Now we’re going to repeatedly sample 50-patients from the population and build up sampling distributions for the mean_thalach and prop_exang point estimates. Take 5000 samples and store the sampling distributions in a variable rep_sample_50.
[Hint - again, use rep_sample_n() to do this.]

```{r}
heart_sample_50 <- heart_data %>%
  rep_sample_n(size = 50, reps = 5000) %>%
  summarise(mean_thalach = mean(thalach),
            prop_exang = sum(exang == 1) / n())
```

8. Visualise the sampling distributions of mean_thalach and prop_exang as histograms. How normal do the sampling distributions look?
```{r}
heart_sample_50 %>%
  ggplot(aes(x = mean_thalach)) +
  geom_histogram()

heart_sample_50 %>%
  ggplot(aes(x = prop_exang)) +
  geom_histogram()

# very normal
```

9. Obtain standard errors of the sampling distribution of mean_thalach and prop_exang directly as the standard deviations of the sampling distributions held in rep_sample_50
```{r}
heart_sample_50 %>%
  summarise(sd_mean_thalach = sd(mean_thalach),
            sd_prop_exang = sd(prop_exang))
```








    Repeat steps 7. and 9. (omitting step 8.) once more with samples of size 100-patients, and again with samples of size 200-patients. How do the standard errors of mean_thalach and prop_exang change as sample size increases? Does the uncertainty in the calculated statistics increase or decrease as sample size increases?
    [Hint - it’s OK to copy, paste and alter code to do this!]
