---
title: "R Notebook"
output: html_notebook
---

```{r}
house_data <- read_csv("ames.csv") %>%
  clean_names()
```

We are going to use a dataset on house sales in Ames, Iowa over a number of years. We will be interested in the SalePrice of houses (though we should clean_names() the data, so this will end up as sale_price). Load the data and explore it.
```{r}
glimpse(house_data)
```

q2. Plot a histogram of sale_price. Is the distribution normal?

```{r}
house_data %>%
  select(sale_price) %>%
  ggplot() +
  geom_histogram(aes(x = sale_price))

# is it normal? No, it is right-skewed.
```

q3. Calculate the population parameter mean of sale_price. Store it in a variable, as we’ll need this value later when we check if each CI contains the parameter value or not.
```{r}
mean_sale_price <- house_data %>%
  select(sale_price) %>%
  summarise(mean = mean(sale_price))
```

q4. take 1000 samples each of size 100 from sale_price
```{r}

results <- tibble()
for (sample_num in 1:1000){

  # get a new 100-observation random sample from ames
  this_sample <- house_data %>%
    rep_sample_n(size = 100, reps = 1) %>%
    ungroup() %>%
    select(sale_price)
  
  # create a bootstrap distribution of mean(sale_price)
  # use 1000 reps
  this_bootstrap <- this_sample %>%
    specify(response = sale_price) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "mean")

  # get 95% CI for this bootstrap distribution
  this_ci <- this_bootstrap %>%
    get_ci(level = 0.95, type = "percentile")
  
  # prepare new row for results, storing CI bounds
  this_result <- c(
    sample_num = sample_num, 
    lower = this_ci %>% 
      pull(lower_ci), 
    upper = this_ci %>%
      pull(upper_ci)
  )
  
  # add this_result to growing tibble of results
  results <- results %>%
    bind_rows(this_result)
}

```


q5. Pipe results to mutate() to add a new variable popn_param_in_ci which will be TRUE if the population parameter we calculated earlier is between lower and upper, and FALSE otherwise.

[Hint if you stored your parameter in a tibble earlier (e.g. as the result of a summarise() pipe), extract it here as a numeric before you compare with upper and lower]
```{r}
pull(mean_sale_price)

results <- results %>%
  mutate(popn_param_in_ci = 180796.1 < upper & 180796.1 > lower)
```

Q6. Perform an appropriate summarise() to determine the proportion of CIs that contain the population parameter. Our confidence level was 95% (equivalent to a proportion of 0.95). Is your proportion close to that value?
```{r}
results %>%
  summarise(proportion = sum(popn_param_in_ci == TRUE) / n())

# it is close to, but slightly under
```

```{r}

```



2.7 Question 7

Have a think about what all this means, make a list of any points you don’t understand. - Imagine you had only a single 100-observation sample of the ames data (say from a 100 house survey). - Would you know ahead of time whether the CI for mean(sale_price) from your particular sample contains the correct population parameter or not?
