---
title: "R Notebook"
output: html_notebook
---

Week6 day 3

Sampling distribution

```{r}
library(tidyverse)
library(janitor)
```

```{r}
telecomms_data <- read_csv("telecomms_churn.csv") %>%
  clean_names()
```

```{r}
glimpse(telecomms_data)
```

this is what we want to be able to infer from our sample
```{r}
summary_pop <- telecomms_data %>%
  summarise(
    mean_monthly_charges = mean(MonthlyCharges),
    mean_tenure = mean(tenure),
    prop_churn = sum(Churn == "Yes") / n()
  )
```

```{r}
telecomms_data %>%
  ggplot() +
  aes(x = monthly_charges) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7)
```


```{r}
telecomms_data %>%
  ggplot() +
  aes(x = tenure) +
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7)
```


```{r}
telecomms_data %>%
  ggplot() +
  aes(x = churn) +
  geom_bar()
```

```{r}
library(infer)
```

creating sample 
```{r}
sample_200 <- telecomms_data %>%
  rep_sample_n(size = 200, reps = 1)
```

```{r}
groups(sample_200)
```

same calc for sample
```{r}
summary_sample <- sample_200 %>%
  ungroup() %>%
  summarise(mean_monthly_charge = mean(monthly_charges),
            mean_tenture = mean(tenure),
            prop_churn = sum(churn == "Yes") / n()
  )
```
- sample error
= point est - pop parameter

```{r}
summary_sample - summary_pop
```

central limit theorem
implies that even when a model is not normally distributed, because a normal
distribution is created, probabilistic and statistical models that apply to normal
distributions can also apply to non normal distributions


```{r}
rep_sample_200 <- telecomms_data %>%
  rep_sample_n(size = 200, reps = 5000) %>%
  summarise(
    mean_monthly_charge = mean(monthly_charges),
            mean_tenure = mean(tenure),
            prop_churn = sum(churn == "Yes") / n()
  )
```

```{r}
monthly_charges_plot <- rep_sample_200 %>%
  ggplot(aes(x = mean_monthly_charge)) +
  geom_histogram(col = "white", fill = "steel blue")
```

task
```{r}
mean_tenure_plot <- rep_sample_200 %>%
  ggplot(aes(x = mean_tenure)) +
  geom_histogram(col = "white", fill = "steel blue")


prop_churn_plot <- rep_sample_200 %>%
  ggplot(aes(x = prop_churn)) +
  geom_bar(col = "white", fill = "steel blue")
```

standard error
- spread of sampling distribution
- wide spread denotes a high degree of uncertainty





