---
title: "R Notebook"
output: html_notebook
---

Finishing lesson on two sample hypothesis 
Paired / dependent sample test

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

```{r}
corfu <- read_csv("corfu.csv")
florence <- read_csv("florence.csv")
```

null hypothesis: the price of lets in florence is the same as those in corfu
alt hypothesis: the price of lets in florence is lower than those in corfu

sig level: .05

```{r}
# creating one dataframe to work with
apart_prices <- bind_rows(florence, corfu, .id = "location") %>%
  mutate(location = if_else(location == "1",
                            "florence",
                            "corfu")) %>%
  select(-...1)

```

```{r}
apart_prices %>%
  ggplot(aes(x = location, y = price)) +
  geom_boxplot()
```

```{r}
null_distr <- apart_prices %>%
  specify(price ~ location) %>%
  hypothesise(null = "independence") %>%
  generate(reps = 10000, type = "permute") %>%
  calculate(stat = "diff in means",
            order = c("florence", "corfu"))
  
```

```{r}
obs_stat_apart <- apart_prices %>%
  specify(price ~ location) %>%
  calculate(stat = "diff in means",
            order = c("florence", "corfu"))
```

```{r}
null_distr %>%
  visualise() +
  shade_pvalue(obs_stat = obs_stat_apart,
                direction = "left")
```

```{r}
p_value_apart <- null_distr %>%
  get_p_value(obs_stat = obs_stat_apart,
              direction = "left")

p_value_apart
```
2 sample tests
Paired aka dependent samples

```{r}
books <- read_csv("ucla_textbooks_f18.csv")

```


Is there a significant difference between the new price on campus, and an amazon?

```{r}
books_diff <- books %>%
  mutate(diff_new = bookstore_new - amazon_new) %>%
  filter(!is.na(diff_new))

books_diff %>%
  ggplot(aes(x = diff_new)) +
  geom_histogram(col = "white")
```

Hypotheses:
H0: mu(diff_new) = 0
Ha: mu(diff_new) != 0

```{r}
null_distribution <- books_diff %>%
  specify(response = diff_new) %>%
  hypothesise(null = "point", mu = 0) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")
```

```{r}
observed_stat <- books_diff %>%
  specify(response = diff_new) %>%
  calculate(stat = "mean")

observed_stat
```

```{r}
null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat,
                direction = "both")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat,
              direction = "both")
p_value
```

With proportion
two sample testing, testing difference in proportions

Testing for whether the proportion of women who voted for Obama is significantly 
differnet to those who voted for romney
```{r}
poll <- read_csv("election_data.csv")
```

```{r}
ggplot(poll, aes(x = vote, fill = gender)) +
  geom_bar(position = "fill") +
  coord_flip()
```


```{r}
obs_stat_poll <- poll %>%
  specify(gender ~ vote,
          success = "woman") %>%
  calculate(stat = "diff in props",
            order = c("obama", "romney"))
```

```{r}
null_dist <- poll %>%
  specify(gender ~ vote,
          success = "woman") %>%
  hypothesise(null = "independence") %>%
  generate(reps = 5000, type = "permute") %>%
  calculate(stat = "diff in props",
            order = c("obama", "romney"))

head(null_dist)
```

```{r}
null_dist %>%
  visualise() +
  shade_p_value(obs_stat = obs_stat_poll,
                direction = "both")
```

```{r}
p_value_poll <- null_dist %>%
  get_p_value(obs_stat = obs_stat_poll,
              direction = "both")

p_value_poll
```

