---
title: "R Notebook"
output: html_notebook
---

Week7 day 2 - Spatial Data / spatial analysis

What do you think spatial data looks like?
  Latitude and longitude
  
  
What do you think spatial data visualisation looks like?
  
Spatial (geospatial data) is data about geographical locations

## How is spatial data represented?


Spatial data are special - there are meaning to the points; while encoded
as numbers (dbl's), these numbers actually have a special meaning

Spatial vectors (non-R specific)

Spatial vectors encode spatial data - there are 3 main types

- point (1, 4)
- line ((1,4), (4,5))
- polygon ((2,2), (4,2), (5,3), etc...)

point; train stations, wells, etc..
lines; roads..
polygons; council areas, lakes

## Spatial data in R

Sometimes we just get given data frames with lat/long columns

Another option is to load in shapefiles

to do this, we use the package sf - simplefeatures

```{r}
library(sf)
```

a feature = a geometry

```{r}
north_carolina <- st_read(system.file("shape/nc.shp", package = "sf"))
```

```{r}
library(tidyverse)
```


```{r}
north_carolina %>%
  as_tibble()
```

```{r}
nc_geo <- st_geometry(north_carolina)

nc_geo[[1]]
```

```{r}
class(north_carolina)
```
it's a dataframe with geometries


## Plotting geometries


```{r}
plot(north_carolina["AREA"])
```

```{r}
# first row, "AREA" column
plot(north_carolina[1, "AREA"])
```

Task - Have a look through some of the variables within your north_carolina dataset, and see if you can create a spatial plot using the techniques above.
```{r}
plot(north_carolina["FIPS"])
```
## ggplot and sf

ggplot can plot spatial with sf, called geom_sf

```{r}
north_carolina %>%
  ggplot() +
  geom_sf(aes(fill = SID74), colour = "black") +
  theme_bw()
```

Task - try plotting another feature for NC. What does it tell you? play around
with ggplot aspects (e.g. colour)
```{r}
north_carolina %>%
  ggplot() +
  geom_sf(aes(fill = FIPS), colour = "black", show.legend = FALSE) +
  theme_void() +
  ylim(20, 50)
  
```

```{r}
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
```

Using rnaturalearth package we can import boundaries of countries at various
levels

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
world %>%
  as_tibble() %>%
  head(5)
```

So we've got the data, now we can plot it

```{r}
world %>%
  ggplot() +
  geom_sf() +
  labs(x = "longitude", y = "latitude", title = "World Map")
```

We can plot actual data now: 

```{r}
world %>%
  ggplot() +
  geom_sf(aes(fill = pop_est)) +
  scale_fill_viridis_c(trans = "sqrt")
```

Task: Recap your knowledge from ggplot week, and set your geom_sf aesthetic to be filled with the estimated gdp (gdp_md_est variable). Extra points if you make your map colour blind friendly!
What does your plot tell you? What does it tell you compared to the population?

```{r}
world %>%
  ggplot() +
  geom_sf(aes(fill = gdp_md_est)) +
  scale_fill_continuous(type = "viridis") +
  theme_void()

# It tells me that while China and India are greater than the rest of the world
# in terms of population, the US is greater than the China and India in terms
# of GDP (despite having lower population)
```

With our world data frame, we can filter for specific countries, using the 
tidyverse

```{r}
country_italy <- world %>%
  filter(name == "Italy")
```

```{r}
# plot just italy

country_italy %>%
  ggplot() +
  geom_sf() +
  labs(x = "Longitude",
       y = "Latitude",
       title = "Italy")
```


```{r}
country_denmark <- world %>%
  filter(name == "Denmark")

country_denmark %>%
  ggplot() +
  geom_sf() +
  labs(x = "Longitude",
       y = "Latitude",
       title = "Danmark")

country_faroes <- world %>%
  filter(name == "Faeroe Is.")

country_faroes%>%
  ggplot() +
  geom_sf() +
  labs(x = "Longitude",
       y = "Latitude",
       title = "Færøerne")

```

## Zooming in on particular parts of the world

- we can subset our graph by limiting the x and y range using coord_sf()

```{r}
world %>%
  ggplot() +
  geom_sf() +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

Add informative labels

- we need to tell ggplot where to put the label
- therefore we need to calculate where we want to put the labels (centre of 
each feature)
- we need to calculate the centres (centroids)

```{r}
world %>%
  mutate(centre = st_centroid(st_make_valid(geometry))) %>%
  as_tibble() %>%
  select(name, centre)
```

```{r}
world_with_centres <- world %>%
  mutate(centre = st_centroid(st_make_valid(geometry))) %>%
  mutate(lat = st_coordinates(centre)[,1],
         long = st_coordinates(centre)[,2])
```


```{r}
# centre geometries
# purely illustrative

centre_geo <- world %>%
  mutate(centre = st_centroid(st_make_valid(geometry))) %>%
  select(centre)
```


```{r}
centre_geo[[1]][[1]]
```


```{r}
world_with_centres %>%
  ggplot() +
  geom_sf() +
  geom_text(aes(x = lat, y = long, label = name), colour = "darkblue",
            fontface = "bold", check_overlap = TRUE, size = 3) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

We can add additional information using annotate 

- add one that says: Gulf of Mexico

```{r}
world_with_centres %>%
  ggplot() +
  geom_sf() +
  geom_text(aes(x = lat, y = long, label = name), colour = "darkblue",
            fontface = "bold", check_overlap = TRUE, size = 3) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", size = 5,
           fontface = "italic")+
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) 
```

Task
```{r}
world_with_centres %>%
  ggplot() +
  geom_sf(aes(fill = income_grp)) +
  geom_text(aes(x = lat, y = long, label = name), colour = "black",
            fontface = "bold", check_overlap = TRUE, size = 3) +
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", size = 3,
           fontface = "italic")+
  coord_sf(xlim = c(-110.15, -60.12), ylim = c(2.65, 50.97), expand = TRUE) +
  labs(x = "Latitude",
       y = "Longitude",
       fill = "OECD Income group")
```

## Interactive maps with leaflet

```{r}
library(leaflet)
```


```{r}
leaflet() %>%
  addTiles() %>% #basemap
  addMarkers(lng = 174.768, lat = -36.852, popup = "The birthplace of R")
```

That was our (maybe) first leaflet

Get some spatial data from the web
- turn it into a format R can work with
- visualise it using leaflet

```{r}
library(jsonlite)
```

```{r}
colorado_data_url <-
  "https://data.colorado.gov/resource/j5pc-4t32.json?&county=BOULDER"
```

```{r}
head(readLines(colorado_data_url))
```

jsonlite package has a lot of functions to help work with json data

json is like a list of lists in R - sometimes we need to recursively sift through
to extract the relevant data

```{r}
colorado_water <- fromJSON(colorado_data_url) %>%
  jsonlite::flatten(recursive = TRUE)
```

```{r}
# wrangling
colorado_water_clean <- colorado_water %>%
  select(-location.needs_recoding) %>%
  mutate(across(c(starts_with("location"), amount), as.numeric)) %>%
  filter(!is.na(location.latitude), !is.na(location.longitude))
  
```

```{r}
# visualise with leaflet

colorado_water_clean %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = ~location.longitude,
                   lat = ~location.latitude)
```

- incidences of surface water in and around Boulder, Colorado

```{r}
colorado_water_clean %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = ~location.longitude,
                   lat = ~location.latitude,
                   radius = ~log(amount), weight = 1)
```

## Clustering

Let's have a look at what addMarkers looks like if we have lots of points

```{r}
colorado_water_clean %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lng = ~location.longitude,
                   lat = ~location.latitude)
```

We can add clustering options

```{r}
colorado_water_clean %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lng = ~location.longitude,
                   lat = ~location.latitude,
             clusterOptions = markerClusterOptions())
```

## Leaflet in shiny

```{r}
whisky_df <- CodeClanData::whisky %>%
  rename(long = Latitude,
         lat = Longitude)
```

maybe too many?
allow user to filter for specific regions
only view distilleries in that region

```{r}
whisky_df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lat = ~lat, lng = ~long, popup = ~Distillery)
```

